---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(insidertrade.cache = TRUE)
```

# insidertrade

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

insidertrade provides access to SEC insider trading data (Forms 3/4/5) from two sources:

- [Insider transactions bulk data sets](https://www.sec.gov/data-research/sec-markets-data/insider-transactions-data-sets) (quarterly TSV downloads)
- [EDGAR API](https://www.sec.gov/search-filings/edgar-application-programming-interfaces) (company filing metadata and Form 4 XML transaction details)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/insidertrade")
```

## Configuration

The SEC requires a valid User-Agent with contact information for all API requests.
Set this once per session before making any calls:

```{r}
#| eval: false
options(insidertrade.user_agent = "your@email.com")
```

Optionally, enable caching to avoid re-downloading data:

```{r}
#| eval: false
options(insidertrade.cache = TRUE)
```

## Usage

```{r}
#| label: setup
#| message: false
library(data.table)
library(insidertrade)
```

### Bulk data

Download and parse all Form 3/4/5 tables for Q4 2025:

```{r}
#| label: bulk
trans <- sec_transactions(2025, 4)
```

### Open-market purchases

Filter to open-market purchases (code `"P"`) by officers and directors — the most
informative signal for insider sentiment:

```{r}
#| label: buys
buys <- trans[
  trans_code == "P" & grepl("Officer|Director", rptowner_relationship)
]
buys[, let(value = trans_shares * trans_pricepershare)]
```

### Top insider buys

Rank companies by the number of distinct insider buyers:

```{r}
#| label: top-buys
top <- buys[,
  .(n_insiders = uniqueN(rptownercik), total_usd = sum(value, na.rm = TRUE)),
  by = .(ticker = issuertradingsymbol, company = issuername)
]
setorder(top, -n_insiders, -total_usd)
head(top, 10)
```

### Cluster buys

Identify stocks where 3 or more insiders bought within the quarter — a strong
bullish signal:

```{r}
#| label: cluster
cluster <- buys[,
  .(
    n_insiders = uniqueN(rptownercik),
    total_usd = sum(value, na.rm = TRUE),
    first_buy = min(trans_date),
    last_buy = max(trans_date)
  ),
  by = .(ticker = issuertradingsymbol, company = issuername)
][n_insiders >= 3L]
setorder(cluster, -n_insiders, -total_usd)
head(cluster, 10)
```

### Buy/sell ratio

Compute the market-wide buy/sell ratio by month — a classic contrarian indicator.
High sell ratios historically correlate with market tops:

```{r}
#| label: buy-sell-ratio
#| dpi: 300
library(ggplot2)

open_market <- trans[trans_code %in% c("P", "S")]
open_market[, let(month = as.Date(format(trans_date, "%Y-%m-01")))]
ratio <- open_market[, .(
  n_buys = sum(trans_code == "P"),
  n_sells = sum(trans_code == "S"),
  buy_usd = sum(fifelse(trans_code == "P", trans_shares * trans_pricepershare, 0), na.rm = TRUE),
  sell_usd = sum(fifelse(trans_code == "S", trans_shares * trans_pricepershare, 0), na.rm = TRUE)
), by = month]
ratio[, let(sell_buy_ratio = sell_usd / buy_usd)]

ggplot(ratio, aes(x = month, y = sell_buy_ratio)) +
  geom_col() +
  scale_x_date(date_labels = "%b %Y") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.2),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_blank()
  ) +
  labs(title = "Insider sell/buy ratio by dollar volume (Q4 2025)")
```

### Daily purchase activity

```{r}
#| label: daily
#| dpi: 300
daily <- buys[, .(n_purchases = .N, total_usd = sum(value, na.rm = TRUE)), by = trans_date]

ggplot(daily, aes(x = trans_date, y = n_purchases)) +
  geom_col() +
  scale_x_date(date_labels = "%b %d") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.2),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_blank()
  ) +
  labs(title = "Daily insider purchases by officers and directors (Q4 2025)")
```

### EDGAR API

Look up a company by ticker and fetch their insider filings:

```{r}
#| label: edgar
tickers <- sec_tickers()
cik <- tickers[ticker == "AAPL", cik]
filings <- edgar_insider_filings(cik)
filings
```

### Form 4 transaction details

Parse the Form 4 XML to see the actual transactions — what was traded, how many
shares, and at what price:

```{r}
#| label: form4
form4s <- filings[form == "4"]
txns <- rbindlist(lapply(seq_len(min(10L, nrow(form4s))), function(i) {
  edgar_form4(cik, form4s$accessionnumber[i], form4s$primarydocument[i])
}))
txns[, .(owner_name, transaction_date, transaction_code, security_title, shares, price_per_share)]
```

## Related work

* [insiderTrades](https://github.com/US-Department-of-the-Treasury/insiderTrades)
* [finreportr](https://github.com/sewardlee337/finreportr)
* [edgarWebR](https://github.com/mwaldstein/edgarWebR)
* [finstr](https://github.com/bergant/finstr)
* [tidyedgar](https://cran.r-project.org/package=tidyedgar)
* [XBRL](https://cran.r-project.org/package=XBRL)
