---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# insidertrade

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

insidertrade provides access to SEC insider trading data (Forms 3/4/5) from two sources:

- [Insider transactions bulk data sets](https://www.sec.gov/data-research/sec-markets-data/insider-transactions-data-sets) (quarterly TSV downloads)
- [EDGAR submissions API](https://www.sec.gov/search-filings/edgar-application-programming-interfaces) (company filing metadata)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/insidertrade")
```

## Usage

```{r}
#| label: demo
#| message: false
#| dpi: 300
library(data.table)
library(insidertrade)

# download and parse all Form 3/4/5 tables for Q2 2024
data <- sec_form345(2024, 2)
str(lapply(data, dim))

# get transactions joined with submission and owner details
trans <- sec_transactions(2024, 2)

# filter to open-market purchases by officers and directors
buys <- trans[
  trans_code == "P" &
    grepl("Officer|Director", rptowner_relationship) &
    trans_date >= "2024-04-01" &
    trans_date <= "2024-06-30"
]

# top 10 companies by number of distinct insider buyers
top <- buys[,
  .(
    n_insiders = uniqueN(rptownercik),
    total_shares = sum(trans_shares, na.rm = TRUE)
  ),
  by = .(ticker = issuertradingsymbol, company = issuername)
][order(-n_insiders, -total_shares)]
head(top, 10)

# plot insider purchases over the quarter
library(ggplot2)

daily <- buys[,
  .(n_purchases = .N, total_shares = sum(trans_shares, na.rm = TRUE)),
  by = trans_date
]

ggplot(daily, aes(x = trans_date, y = n_purchases)) +
  geom_col() +
  scale_x_date(date_labels = "%b %d") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.2),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_blank()
  ) +
  labs(title = "Daily Insider Purchases (Q2 2024)")
```

## Related work

* [insiderTrades](https://github.com/US-Department-of-the-Treasury/insiderTrades)
* [finreportr](https://github.com/sewardlee337/finreportr)
* [edgarWebR](https://github.com/mwaldstein/edgarWebR)
* [finstr](https://github.com/bergant/finstr)
* [tidyedgar](https://cran.r-project.org/package=tidyedgar)
* [XBRL](https://cran.r-project.org/package=XBRL)
