---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(insidertrade.cache = TRUE)
```

# insidertrade

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

insidertrade provides access to SEC insider trading data (Forms 3/4/5) from two sources:

- [Insider transactions bulk data sets](https://www.sec.gov/data-research/sec-markets-data/insider-transactions-data-sets) (quarterly TSV downloads)
- [EDGAR submissions API](https://www.sec.gov/search-filings/edgar-application-programming-interfaces) (company filing metadata)

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("pak")
pak::pak("m-muecke/insidertrade")
```

## Usage

```{r}
#| label: demo
#| message: false
#| dpi: 300
library(data.table)
library(insidertrade)

# download and parse all Form 3/4/5 tables for Q2 2024
data <- sec_form345(2025, 4)
str(lapply(data, dim))

# get transactions joined with submission and owner details
trans <- sec_transactions(2025, 4)

# filter to open-market purchases by officers and directors
buys <- trans[
  trans_code == "P" &
    grepl("Officer|Director", rptowner_relationship) &
    trans_date >= "2025-10-01" &
    trans_date <= "2025-12-31"
]

# top 10 companies by number of distinct insider buyers
buys[, value := trans_shares * trans_pricepershare]
top <- buys[,
  .(n_insiders = uniqueN(rptownercik), total_usd = sum(value, na.rm = TRUE)),
  by = .(ticker = issuertradingsymbol, company = issuername)
]
setorder(top, -n_insiders, -total_usd)
head(top, 10)

# plot insider purchases over the quarter
library(ggplot2)

daily <- buys[,
  .(n_purchases = .N, total_shares = sum(trans_shares, na.rm = TRUE)),
  by = trans_date
]

ggplot(daily, aes(x = trans_date, y = n_purchases)) +
  geom_col() +
  scale_x_date(date_labels = "%b %d") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.2),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_blank()
  ) +
  labs(title = "Daily Insider Purchases (Q4 2025)")
```

You can also look up companies by ticker and fetch their insider filings from EDGAR:

```{r}
#| label: edgar
tickers <- sec_tickers()
cik <- tickers[ticker == "AAPL", cik]
filings <- edgar_insider_filings(cik)
filings
```

To see the actual transaction details, parse a Form 4 filing:

```{r}
#| label: form4
form4s <- filings[form == "4"]
txns <- rbindlist(lapply(seq_len(min(5L, nrow(form4s))), function(i) {
  edgar_form4(cik, form4s$accessionnumber[i], form4s$primarydocument[i])
}))
txns[, .(owner_name, transaction_date, transaction_code, security_title, shares, price_per_share)]
```

## Related work

* [insiderTrades](https://github.com/US-Department-of-the-Treasury/insiderTrades)
* [finreportr](https://github.com/sewardlee337/finreportr)
* [edgarWebR](https://github.com/mwaldstein/edgarWebR)
* [finstr](https://github.com/bergant/finstr)
* [tidyedgar](https://cran.r-project.org/package=tidyedgar)
* [XBRL](https://cran.r-project.org/package=XBRL)
